Compositional equivalences based on Open pNets\thanks{This work was partially funded by the Associated Team FM4CPS between INRIA and ECNU, Shanghai}

### 2019-07-17 2016-Weak-FH-Bisim/WeakBisim.tex
Added TODOs for the example

\subsection{Running Example}
To illustrate this work, we use the well-known Alternate bit protocol
(ABP) \TODO{Ref ?}, that provides safe transport of data between two processes,
over unsafe media. This protocol has been described and studied using
many approaches, because it is small enough, and can be easily
verified using exact finite models. Our version here is more
ambitious, encoding both the data and the control bits in a symbolic
way. One can imagine that this could easily be extended using bounded
integers, as in the sliding window protocol, or with more complex data
structures, in order to verify realistic protocols.

@@

%%\TODO{rewrite example with new OTs}
\begin{example}\emph{An open-transition.}
  \label{OT:ABP}
At the upper level, the $ABP$ pNet has 2 holes and $Protocol$ has
a subnet. One of its possible open transitions (synchronizing the hole $P$
with the \emph{Protocol}) is:

@@

\LUDO{Je sais pas de quoi parle le prochain todo :-( l'exemple?}
\TODO{Ceci n'est pas nouveau, n'est-ce pas, est-ce utile pour le lecteur ici ? Ou est-ce qu'on bouge tous ces ``details techniques'' dans une annexe ? Si oui, je refais avec un morceau de ABP}
\begin{example} \emph{Using the operational rules to compute
    open-transitions}
  In Fig. \ref{usingrules:OT2} we show the deduction tree used to construct and prove the 
  open transition $OT_2$ of \texttt{EnableCompL} (see example page \pageref{OT:ABP-composed}).
  The rule uses \TrUn\ for the $\delta$ transition of $C_3$, for the $l$ transition of $C_4$, then combines the result using the $a_4$ vector of the bottom pNet node, and the $\underline{\delta(x)}$ vector of the top node.

\begin{figure}[h]
\begin{mathpar}
  \small
  \inferrule
    {\inferrule
        {0 \xrightarrow {\delta}_{C_3} 1}
        {C_3
          \models
            \openrule{
              0 \xrightarrow {\delta}_{C_3} 1,\,
              \{\xrightarrow{\delta(x_1)}_P\},\,
              v_1=\delta(x_1)}
                      {\ostate{0}\OTarrow{v_1}\ostate{1}}
        }\\
%      \sm{fresh}v         \\
      \inferrule%*[right={L_1}]
        {
%          \sm{fresh}{a_Q} \\
          \inferrule
              {0 \xrightarrow{l}_{C_4} 0}
              % {\ostate{0}\xrightarrow{l}\ostate{0}}
              {C_4
                \models
                \openrule{
                      0 \xrightarrow l_{C_4} 0,\,
                      \sm{Pred}_{C_4}}
                      {\ostate{0}\OTarrow{l}\ostate{0}}
              }
        }
        {
          \textrm{Q>>R}\models
              \openrule
                  { 0 \xrightarrow {l}_{C_4} 0,\,
                    \{\xrightarrow{acc(x_2)}_Q\},\,
                    \ v_2=acc(x_2)
                  }
                  {\ostate{0}\OTarrow{v_2}\ostate{0}}
        }   
    }
    {
     \textrm{P>>(Q>>R)}
     \models
     \openrule
         { 0 \xrightarrow{\delta}_{C_3} 1, \\ 0 \xrightarrow{l}_{C_4} 0,\\
           \{\xrightarrow{\delta(x)}_P,\,\xrightarrow{acc(x)}_Q\}, \\
            a_3=v_1 \wedge v=a_3 \wedge x_1=x_2
           }
         {\ostate{00} \OTarrow{v} \ostate{10}}
      }\vspace{-4ex}
\end{mathpar}
  \caption{Proof of transition $OT_2$ (with interaction of processes $P$ and $Q$) for 
  ``P>>(Q>>R)'' \TODO{Change for the new example or remove?}}
  \label{usingrules:OT2}
\end{figure}

\paragraph{Variable management.}
The variables in each synchronisation vector are considered local:
for a given pNet expression, we must have fresh local variables for
each occurrence of a vector (i.e. each time we instantiate rule
\TrDeux). Similarly the state variables of each copy of a
given pLTS in the system, must be distinct, and those created for each
application of \TrDeux\ have to be fresh and all distinct. 
This will be implemented within the open-automaton generation algorithm,
e.g. using name generation using a global counter as a suffix.

@@

\TODO{Eric: full (???) result of the tool execution on the running example}

@@

%% \TODO{ the formal view is not much interesting}

%% \begin{property}[Weak open transition transitivity]
%% Let $A$ be an open automaton, and $<J,\mathcal{S},s_0,\WT>$ a weak open automaton derived from $A$.
%% Suppose we have the following weak open transitions in $\WT$:
%% \[\openrule
%%          {
%%            \set{\gamma_1},\Pred_1,\Post_1   }
%%          {s \OTWeakarrow {\tau} s_1}
%% \qquad 
%% \openrule
%%          {
%%            \set{\gamma_2},\Pred_2,\Post_2   }
%%          {s_1 \OTWeakarrow {\alpha} s_2}
%% \qquad
%% \openrule
%%          {
%%            \set{\gamma_3},\Pred_3,\Post_3   }
%%          {s_2 \OTWeakarrow {\tau} s_3}
%% \]
%% Then there is a weak  open transition of the following form in $\WT$:
%% \[\openrule
%%          {
%%            \set{\gamma},\Pred,\Post   }
%%          {s \OTWeakarrow {\alpha} s_3}
%% \]
%% Suppose additionally $\Post_3=({x_k\gets e_k})^{k\in K}$, we have the following
%% \[
%% \begin{array}{rcl}
%% \gamma&=& \gamma_1\dotcup\gamma_2\dotcup\gamma_3\\
%% \Pred&=&\Pred_1\land\Pred_2\subst{\Post_1}\land (\Pred_3\subst{\Post_2})\subst{\Post_1}\\
%% \Post &=& ({x_k\gets ((e_k\subst{\Post_2})\subst{\Post_1})})^{k\in K}
%% \end{array}\]
%% \end{property}

@@

\subsection{Running example}
- Detail of a $\tau^*$ sequence, \\
- construction of the predicate\\
- defn / elimination of the intermediate variables\\
- construction of the weak OT.

>>>>

\subsection{Running Example}
\TODO{Rabea: trouver reference !!}
\TODO{refaire les commentaires}
To illustrate this work, we use the well-known Alternate bit protocol
(ABP) \TODO{Ref ?}, that provides safe transport of data between two processes,
over unsafe media. This protocol has been described and studied using
many approaches, because it is small enough, and can be easily
verified using exact finite models. Our version here is more
ambitious, encoding both the data and the control bits in a symbolic
way. One can imagine that this could easily be extended using bounded
integers, as in the sliding window protocol, or with more complex data
structures, in order to verify realistic protocols.

@@

\TODO{Eric: rewrite example with new Protocol}
\begin{example}\emph{An open-transition.}
  \label{OT:ABP}
At the upper level, the $ABP$ pNet has 2 holes and $Protocol$ has
a subnet. One of its possible open transitions (synchronizing the hole $P$
with the \emph{Protocol}) is:

@@

\TODO{Add a reference to the construction (deduction tree) in the forte paper?}

%% \LUDO{Je sais pas de quoi parle le prochain todo :-( l'exemple?}
%% \TODO{Ceci n'est pas nouveau, n'est-ce pas, est-ce utile pour le lecteur ici ? Ou est-ce qu'on bouge tous ces ``details techniques'' dans une annexe ? Si oui, je refais avec un morceau de ABP}
%% \begin{example} \emph{Using the operational rules to compute
%%     open-transitions}
%%   In Fig. \ref{usingrules:OT2} we show the deduction tree used to construct and prove the 
%%   open transition $OT_2$ of \texttt{EnableCompL} (see example page \pageref{OT:ABP-composed}).
%%   The rule uses \TrUn\ for the $\delta$ transition of $C_3$, for the $l$ transition of $C_4$, then combines the result using the $a_4$ vector of the bottom pNet node, and the $\underline{\delta(x)}$ vector of the top node.

%% \begin{figure}[h]
%% \begin{mathpar}
%%   \small
%%   \inferrule
%%     {\inferrule
%%         {0 \xrightarrow {\delta}_{C_3} 1}
%%         {C_3
%%           \models
%%             \openrule{
%%               0 \xrightarrow {\delta}_{C_3} 1,\,
%%               \{\xrightarrow{\delta(x_1)}_P\},\,
%%               v_1=\delta(x_1)}
%%                       {\ostate{0}\OTarrow{v_1}\ostate{1}}
%%         }\\
%% %      \sm{fresh}v         \\
%%       \inferrule%*[right={L_1}]
%%         {
%% %          \sm{fresh}{a_Q} \\
%%           \inferrule
%%               {0 \xrightarrow{l}_{C_4} 0}
%%               % {\ostate{0}\xrightarrow{l}\ostate{0}}
%%               {C_4
%%                 \models
%%                 \openrule{
%%                       0 \xrightarrow l_{C_4} 0,\,
%%                       \sm{Pred}_{C_4}}
%%                       {\ostate{0}\OTarrow{l}\ostate{0}}
%%               }
%%         }
%%         {
%%           \textrm{Q>>R}\models
%%               \openrule
%%                   { 0 \xrightarrow {l}_{C_4} 0,\,
%%                     \{\xrightarrow{acc(x_2)}_Q\},\,
%%                     \ v_2=acc(x_2)
%%                   }
%%                   {\ostate{0}\OTarrow{v_2}\ostate{0}}
%%         }   
%%     }
%%     {
%%      \textrm{P>>(Q>>R)}
%%      \models
%%      \openrule
%%          { 0 \xrightarrow{\delta}_{C_3} 1, \\ 0 \xrightarrow{l}_{C_4} 0,\\
%%            \{\xrightarrow{\delta(x)}_P,\,\xrightarrow{acc(x)}_Q\}, \\
%%             a_3=v_1 \wedge v=a_3 \wedge x_1=x_2
%%            }
%%          {\ostate{00} \OTarrow{v} \ostate{10}}
%%       }\vspace{-4ex}
%% \end{mathpar}
%%   \caption{Proof of transition $OT_2$ (with interaction of processes $P$ and $Q$) for 
%%   ``P>>(Q>>R)'' \TODO{Change for the new example or remove?}}
%%   \label{usingrules:OT2}
%% \end{figure}

%% \end{example}

\paragraph{Variable management.}
The variables in each synchronisation vector are considered local:
for a given pNet expression, we must have fresh local variables for
each occurrence of a vector (i.e. each time we instantiate rule
\TrDeux). Similarly the state variables of each copy of a
given pLTS in the system, must be distinct, and those created for each
application of \TrDeux\ have to be fresh and all distinct. 
This is implemented within the open-automaton generation algorithm,
using name generation using a global counter as a suffix.

@@

\TODO{Eric: full (???) result of the tool execution on the running example on the Simple Prot Spec, and maybe one example on the Impl. OTs presented in latex readable form.}

@@

\TODO{Eric: try to draw (ATG) the weak transitino versions of the Simple Protocol, both Spec (easy) and Impl (check readability?)}

@@

\subsection{Running example}
- If we already have shown the saturated relation, we only need to argue on bisimulation here.

\TODO{Je ne pense pas que nous ayons un algo de bisim weak operationnel pour la publi, mais ca n'est pas completement impossible nonp[l;us...  Au pire, argumenter intuitivement sur papier, en donnat les classes d'equivalence et les predicats.
    Noter que la version de base est trop simple... Pour avoir des predicats moins triviaux, on peut ajouter un compteur qui compte le nb d'erreurs, et le transmet a la fin}
